---
title: "Лабораторная работа №1"
author: "Гуляев Степан БИСО-01-20"
format: 
  md:
    output-file: README.md
editor: visual
---

## Цель работы

1.  Развить практические навыки использования языка программирования R для обработки данных

2.  Развить навыки работы в Rstudio IDE:

-   Установка пакетов

-   Работа с проектами в Rstudio

-   Настройка и работа с Git

3.  Закрепить знания базовых типов данных языка R и простейших операций с ними

## Задание

Используя программный пакет swirl, освоить базовые операции в языке программирования R.

## Исходные данные

1.  ОС Windows 10
2.  Интерпретатор R v-4.3.1
3.  RStudio Desktop

## Ход работы

### Установка swirl

Введем команды:

-   install.packages("swirl")
-   swirl::swirl()

И будем следовать инструкциям базового курса

### Базовые структурные блоки

#### Шаг 1

In its simplest form, R can be used as an interactive calculator. Type 5 + 7 and press Enter.

```{r}
5 + 7
```

#### Шаг 2

To assign the result of 5 + 7 to a new variable called x, you type x \<- 5 + 7. This can be read as 'x gets 5 plus 7'. Give it a try now.

```{r}
x <- 5 + 7
```

#### Шаг 3

To view the contents of the variable x, just type x and press Enter. Try it now.

```{r}
x
```

#### Шаг 4

Now, store the result of x - 3 in a new variable called y.

```{r}
y <- x - 3
```

#### Шаг 5

What is the value of y? Type y to find out.

```{r}
y
```

#### Шаг 6

Now, let's create a small collection of numbers called a vector. Any object that contains data is called a data structure and numeric vectors are the simplest type of data structure in R. In fact, even a single number is considered a vector of length one.

The easiest way to create a vector is with the c() function, which stands for 'concatenate' or 'combine'. To create a vector containing the numbers 1.1, 9, and 3.14, type c(1.1, 9, 3.14). Try it now and store the result in a variable called z.

```{r}
z <- c(1.1,9,3.14)
```

#### Шаг 7

Anytime you have questions about a particular function, you can access R's built-in help files via the `?` command. For example, if you want more information on the c() function, type ?c without the parentheses that normally follow a function name. Give it a try.

```{r}
?c
```

#### Шаг 8

Type z to view its contents. Notice that there are no commas separating the values in the output.

```{r}
z
```

#### Шаг 9

You can combine vectors to make a new vector. Create a new vector that contains z, 555, then z again in that order. Don't assign this vector to a new variable, so that we can just see the result immediately.

```{r}
c(z,555,z)
```

#### Шаг 10

Numeric vectors can be used in arithmetic expressions. Type the following to see what happens: z \* 2 + 100.

```{r}
z * 2 + 100
```

#### Шаг 11

Take the square root of z - 1 and assign it to a new variable called my_sqrt.

```{r}
my_sqrt <- sqrt(z-1)
```

#### Шаг 12

Print the contents of my_sqrt.

```{r}
my_sqrt
```

#### Шаг 13

Now, create a new variable called my_div that gets the value of z divided by my_sqrt.

```{r}
my_div <- z/my_sqrt
```

#### Шаг 14

Go ahead and print the contents of my_div.

```{r}
my_div
```

#### Шаг 15

To see another example of how this vector 'recycling' works, try adding c(1, 2, 3, 4) and c(0, 10). Don't worry about saving the result in a new variable.

```{r}
c(1,2,3,4) + c(0,10)
```

#### Шаг 16

Try c(1, 2, 3, 4) + c(0, 10, 100) for an example.

```{r}
c(1, 2, 3, 4) + c(0, 10, 100)
```

#### Шаг 17

In many programming environments, the up arrow will cycle through previous commands. Try hitting the up arrow on your keyboard until you get to this command (z \* 2 + 100), then change 100 to 1000 and hit Enter. If the up arrow doesn't work for you, just type the corrected command.

```{r}
z * 2 + 1000
```

#### Шаг 18

You can type the first two letters of the variable name, then hit the Tab key (possibly more than once). Most programming environments will provide a list of variables that you've created that begin with 'my'. This is called auto-completion and can be quite handy when you have many variables in your workspace. Give it a try. (If auto-completion doesn't work for you, just type my_div and press Enter.)

```{r}
my_div
```

### Рабочие пространства и файлы

#### Шаг 1

Determine which directory your R session is using as its current working directory using getwd().

```{r}
getwd()
```

#### Шаг 2

List all the objects in your local workspace using ls().

```{r}
ls()
```

#### Шаг 3

Assign 9 to x using x \<- 9.

```{r}
x <- 9
```

#### Шаг 4

Now take a look at objects that are in your workspace using ls().

```{r}
ls()
```

#### Шаг 5

List all the files in your working directory using list.files() or dir().

```{r}
dir()
```

#### Шаг 6

As we go through this lesson, you should be examining the help page for each new function. Check out the help page for list.files with the command ?list.files.

```{r}
?list.files
```

#### Шаг 7

Use the args() function to determine the arguments to list.files().

```{r}
args(list.files)
```

#### Шаг 8

Use the args() function to determine the arguments to list.files().

```{r}
old.dir <- getwd()
```

#### Шаг 9

Use dir.create() to create a directory in the current working directory called "testdir".

```{r}
dir.create("testdir")
```

#### Шаг 10

Set your working directory to "testdir" with the setwd() command.

```{r}
setwd("testdir")
```

#### Шаг 11

Create a file in your working directory called "mytest.R" using the file.create() function.

```{r}
file.create("mytest.R")
```

#### Шаг 12

This should be the only file in this newly created directory. Let's check this by listing all the files in the current directory.

```{r}
dir()
```

#### Шаг 13

Check to see if "mytest.R" exists in the working directory using the file.exists() function.

```{r}
file.exists("mytest.R")
```

#### Шаг 14

Check to see if "mytest.R" exists in the working directory using the file.exists() function.

```{r}
file.info("mytest.R")
```

#### Шаг 15

Change the name of the file "mytest.R" to "mytest2.R" by using file.rename().

```{r}
file.rename("mytest.R","mytest2.R")
```

#### Шаг 16

Make a copy of "mytest2.R" called "mytest3.R" using file.copy().

```{r}
file.copy("mytest2.R","mytest3.R")
```

#### Шаг 17

Provide the relative path to the file "mytest3.R" by using file.path().

```{r}
file.path("mytest3.R")
```

#### Шаг 18

You can use file.path to construct file and directory paths that are independent of the operating system your R code is running on. Pass 'folder1' and 'folder2' as arguments to file.path to make a platform-independent pathname.

```{r}
file.path('folder1','folder2')
```

#### Шаг 19

Take a look at the documentation for dir.create by entering ?dir.create . Notice the 'recursive' argument. In order to create nested directories, 'recursive' must be set to TRUE.

```{r}
?dir.create
```

#### Шаг 20

Create a directory in the current working directory called "testdir2" and a subdirectory for it called "testdir3", all in one command by using dir.create() and file.path().

```{r}
dir.create(file.path('testdir2','testdir3'),recursive = TRUE)
```

#### Шаг 21

Go back to your original working directory using setwd(). (Recall that we created the variable old.dir with the full path for the orginal working directory at the start of these questions.)

```{r}
setwd(old.dir)
```

### Последовательности чисел

#### Шаг 1

The simplest way to create a sequence of numbers in R is by using the `:` operator. Type 1:20 to see how it works.

```{r}
1:20
```

#### Шаг 2

That gave us every integer between (and including) 1 and 20. We could also use it to create a sequence of real numbers. For example, try pi:10.

```{r}
pi:10
```

#### Шаг 3

That gave us every integer between (and including) 1 and 20. We could also use it to create a sequence of real numbers. For example, try pi:10.

```{r}
pi:10
```

#### Шаг 4

What happens if we do 15:1? Give it a try to find out

```{r}
15:1
```

#### Шаг 5

Pull up the documentation for `:` now.

```{r}
?':'
```

#### Шаг 6

The most basic use of seq() does exactly the same thing as the `:` operator. Try seq(1, 20) to see this.

```{r}
seq(1,20)
```

#### Шаг 7

This gives us the same output as 1:20. However, let's say that instead we want a vector of numbers ranging from 0 to 10, incremented by 0.5. seq(0, 10, by=0.5) does just that. Try it out.

```{r}
seq(0,10,by=0.5)
```

#### Шаг 8

Or maybe we don't care what the increment is and we just want a sequence of 30 numbers between 5 and 10. seq(5, 10, length=30) does the trick. Give it a shot now and store the result in a new variable called my_seq.

```{r}
my_seq <- seq(5, 10, length = 30)
```

#### Шаг 9

To confirm that my_seq has length 30, we can use the length() function. Try it now.

```{r}
length(my_seq)
```

#### Шаг 10

There are several ways we could do this. One possibility is to combine the `:` operator and the length() function like this: 1:length(my_seq). Give that a try.

```{r}
1:length(my_seq)
```

#### Шаг 11

There are several ways we could do this. One possibility is to combine the `:` operator and the length() function like this: 1:length(my_seq). Give that a try.

```{r}
seq(along.with = my_seq)
```

#### Шаг 12

There are several ways we could do this. One possibility is to combine the `:` operator and the length() function like this: 1:length(my_seq). Give that a try.

```{r}
seq_along(my_seq)
```

#### Шаг 13

If we're interested in creating a vector that contains 40 zeros, we can use rep(0, times = 40). Try it out.

```{r}
rep(0, times = 40)
```

#### Шаг 14

If we're interested in creating a vector that contains 40 zeros, we can use rep(0, times = 40). Try it out.

```{r}
rep(c(0,1,2), times = 10)
```

#### Шаг 15

Finally, let's say that rather than repeating the vector (0, 1, 2) over and over again, we want our vector to contain 10 zeros, then 10 ones, then 10 twos. We can do this with the `each` argument. Try rep(c(0, 1, 2),each = 10).

```{r}
rep(c(0,1,2), each = 10)
```

### Векторы

#### Шаг 1

First, create a numeric vector num_vect that contains the values 0.5, 55, -10, and 6.

```{r}
num_vect <- c(0.5, 55, -10, 6)
```

#### Шаг 2

Now, create a variable called tf that gets the result of num_vect < 1, which is read as 'num_vect is less than 1'.

```{r}
tf <- num_vect < 1
```

#### Шаг 3

Print the contents of tf now.

```{r}
tf
```

#### Шаг 4

Print the contents of tf now.

```{r}
num_vect >= 6
```

#### Шаг 5

Create a character vector that contains the following words: "My", "name", "is". Remember to enclose each word in its own set of double quotes, so that R knows they are character strings. Store the vector in a variable called my_char.

```{r}
my_char <- c("My","name","is")
```

#### Шаг 6

Print the contents of my_char to see what it looks like.

```{r}
my_char
```

#### Шаг 7

Print the contents of my_char to see what it looks like.

```{r}
paste(my_char, collapse = " ")
```

#### Шаг 8

To add (or 'concatenate') your name to the end of my_char, use the c() function like this: c(my_char, "your_name_here"). Place your name in double quotes where I've put "your_name_here". Try it now, storing the result in a new variable called my_name.

```{r}
my_name <- c(my_char, "Stepan")
```

#### Шаг 9

To add (or 'concatenate') your name to the end of my_char, use the c() function like this: c(my_char, "your_name_here"). Place your name in double quotes where I've put "your_name_here". Try it now, storing the result in a new variable called my_name.

```{r}
my_name
```

#### Шаг 10

Now, use the paste() function once more to join the words in my_name together into a single character string. Don't forget to say collapse = " "!

```{r}
paste(my_name,collapse = " ")
```

#### Шаг 11

In the simplest case, we can join two character vectors that are each of length 1 (i.e. join two words). Try paste("Hello", "world!", sep = " "), where the `sep` argument tells R that we want to separate the joined elements with a single space.

```{r}
paste("Hello", "world!", sep = " ")
```

#### Шаг 12

In the simplest case, we can join two character vectors that are each of length 1 (i.e. join two words). Try paste("Hello", "world!", sep = " "), where the `sep` argument tells R that we want to separate the joined elements with a single space.

```{r}
paste(c(1:3), c("X","Y","Z"), sep = "")
```

#### Шаг 13

Vector recycling! Try paste(LETTERS, 1:4, sep = "-"), where LETTERS is a predefined variable in R containing a character vector of all 26 letters in the English alphabet.

```{r}
paste(LETTERS,1:4, sep = "-")
```

## Оценка результата

В результате лабораторной работы мы прошли 4 урока по языку программирования R, тем самым разобравшись в базовых аспектах языка.

## Вывод

Мы получили базовые навыки работы с RStudio Desktop и языком R



















